\chapter{Quantum Algorithms}
\label{chap:qalg}

\chapabstract{This chapter applies operational process theories to the study of quantum algorithms in three ways. The first section connects the Fourier Transform to strongly complementary observables in an oPT. This provides a new mathematical setting for Fourier theory and pinpoints the reason why the Fourier Transform appears naturally in quantum computation. These results are ongoing collaborative work by the author and Stefano Gogioso that updates our preprint~\cite{gogioso2015fourier}.

The second section applies oPTs to the verification and generalization of quantum blackbox algorithms. We provide a general construction of unitary oracles in arbitrary process theories and prove an equivalence between the unitarity of these oracles and the existence of complementary observables. We then extend work by Vicary~\cite{vicary-tqa} to present a quantum blackbox algorithm for a new problem that identifies group homomorphisms. These results are extensions of collaborative work by the author and Jamie Vicary published in~\cite{zeng2014abstract}.

The third section of this chapter presents a new toy model for quantum blackbox algorithms in the oPT of sets and relations along with several mathematical results.  This setting is usually interpreted as a model for nondeterministic classical computation and so demonstrates a technique for modelling quantum algorithms classically. Results of this section are adapted from work by the author in the preprint~\cite{zeng2015models}.
}
\input{strongcomplft}
\input{blackbox}
\input{QAlgRel}
